---
title: "Quantifying lexical contact in Daghestan"
author: "M. Daniel, I. Chechuro, S. Verhees"
date: "20.02.2018 <br> Linguistic Convergence Lab Seminar"
output:
  xaringan::moon_reader:
    css: ["default", "my_theme.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning= FALSE,
                      fig.width=10.4,
                      fig.height=7)
library(tidyverse)
library(lingtypology)
library(phangorn)
library(knitr)
```

## Outline of the talk

- Comparison of our shortlist and the WOLD questionnaire

--

- Data from the field: Rutul region

  + Similarity and distance matrices
  + Neighbournet visualization
  
--

- Dictionary data: neighbournet
  
--

- Data from the field: Avar contact zone

--

- Problems

--

- Further plans


---


## WOrld Loan Database project: a brief overview

- [WOLD](http://wold.clld.org/meaning) questionnaire - 1814 items 

--

- The aim of the [WOLD](http://wold.clld.org/meaning) project: to test empirically borrowability of different lexical items cross-linguistically

-- 

- For East Caucasian, [WOLD](http://wold.clld.org/meaning) is represented by two languages in central Daghestan: Archi (Lezgic) and Bezhta (Tsezic)

--

- [WOLD](http://wold.clld.org/meaning) includes all parts of speech (including ?? nouns, ?? verbs, ?? adjectives)

--

- Empirically confirms the lexical borrowability hierarchy, with the noun on top

--

- To quantify lexical contact, we thus decided to focus - at this stage of the project - on nouns

--

- We have collected [WOLD](http://wold.clld.org/meaning) nouns from the dictionaries of all major Daghestanian languages and some of their immediate neighbors (Avar, standard Dargwa, Lezgian, Lak, Kumyk; Chechen and Azerbaijani) - DICT dataset below

---

## Shortlist

- We decided to also compile a short list that can be quickly collected in the field from several speakers - WOLD is too long

--

- The short list was supposed to show a high ratio of horyzontally borrowed items (higher than full WOLD), lower ratio of vertical borrowings

--

- Thus: nouns coming from mid-range of the WOLD list (from 30 to 45) - to avoid conservative lexical items and lexical items typical in vertical borrowings; names of locally non-existent objects were avoided; some assumedly locally important nouns were added

--

- Resulting shortlist: some 225 nouns

---



## Shortlist versus WOLD longlist

```{r, echo = F, results = 'asis'}

count <- read_tsv("loancount.csv")
colnames(count) <- c("Archi", "long", "%", "short", "%",
                     "Bezhta", "long", "%", "short", "%")

kable(count[,1:5], format = "html")


```

Similar percentages for most values, 
*but* a relatively large amount of **non loans**.

---

## Shortlist versus WOLD longlist

```{r, echo = F, results = 'asis'}

count <- read_tsv("loancount.csv")
colnames(count) <- c("Archi", "long", "%", "short", "%",
                     "Bezhta", "long", "%", "short", "%")

kable(count[,6:10], format = "html")


```

Percentage of Georgian borrowings rises relatively.

---

## Shortlist versus WOLD longlist

- The shortlist does not increase the percentage of horyzontal loans; neither does it reduce the percentage of vertical loans - instead, it captures percentages similar to the long list

--

- For Archi, they concur with data on multilingualism (many people speak Avar, some speak Lak)

--

- Relative rise of Georgian borrowings in Bezhta seems to be due to the domain of everyday objects (e.g. 'spoon', 'candle').

---


class: inverse, center, middle

## The rationale for the fieldwork

- Even for minority languages, dictionaries tend to be on the conservative (pure language) side

--

- Collect more naturalistic data on the usage of loanwords from speakers

--

- Collect comparable data from several speakers in a partly controlled experimental setting

--

- Downs: no strict transcription of the data, similarity judgements without etymological investigation

---

## Fieldwork

```{r, fig.height=7}

# general area map

loan <- read_csv("loan_villages.csv")

map.feature(loan$Language,
            features = loan$Language,
            tile = c("Esri.OceanBasemap"),
            color = loan$color,
            latitude = loan$Lat,
            longitude = loan$Lon,
            title = "Language",
            label = loan$EngNames,
            label.hide = F,
            zoom.control = T)

```

---

## Similarity sets

- Lexemes from different lects are grouped into similarity sets
Each set is coded with a number. One language may use several lexical items and thus be ascribed several similarity sets.

--

```{r}

sim <- read_tsv("similarity.csv")
kable(sim, format = "html")

```

--

## From similarity sets to distance matrices

- Convert shared loans into similarity models (neighbornet): Number of times two languages share a similarity set is converted into a distance matrix

- Simplistic method: Count similarities between lects by counting how many similarity sets two lects share. (SV-style below)

- Issue: similarities by cognate are counted (to be excluded?). 

- Issue: NOT having the same loan is counted as similarity, even if different items are used (Are 0's indeed counted?)

- Complification: A similarity sets is only counted if it goes beyond a set of closely related languages.
Lines where similarity sets are contained in e.g. Turkic only or Rutul only are excluded from the counts. (MD-style below)

- Complification: One meaning is counted twice. Two languages are deemed similar within a meaning if the share at least one similarity set for this meaning.
--



---


## Similarity to distance matrix

- The number of times two languages share a similarity set is converted into a distance matrix 

- Distance: the number of shared sets divided by the number of meanings (minus the meanings for which one of the lects had not data)

--

```{r}

dist <- read.csv("rutul_dist.csv", sep = ";")
kable(dist[1:4,1:3], format = "html")

```

--

- Distance matrix forms the input for a neighbournet visualization

---

## Rutul area

```{r}

Rutul_dist <- read.csv("rutul_dist.csv", header = TRUE, sep = ";", na.strings = "NA")
View(Rutul_dist)
Rutul_dist <- as.dist(Rutul_dist)
nnet <- neighborNet(Rutul_dist)
plot(nnet, "2D")

```
---

## Distance between major languages

```{r}

Big_dist <- read.csv("Big_dist.csv", header = TRUE, sep = ";", na.strings = "NA")
View(Big_dist)
Big_dist <- as.dist(Big_dist)
nnet <- neighborNet(Big_dist)
plot(nnet, "2D")

```

---

## Avar area

- Archi (Lezgic) | Bezhta (Tsezic) | Andi (Andic; Zilo dialect), Botlikh (Andic)

--

- All speak Avar as second language. Archi also has Lak influence, Bezhta has Georgian influence. Zilo Andi and Botlikh: only Avar.

--

- Remove borrowings from Arabic/Persian through Avar.

--

- Only **one** word is borrowed in all four languages: *Êalbac'*, 'lion'


---

## Problems


---

## Further plans


---

## Thank you!

We used

[**lingtypology**](https://ropensci.github.io/lingtypology/) and [**xaringan**](https://github.com/yihui/xaringan) for this presentation.
