---
title: "Quantifying lexical contact in Daghestan"
author: "M. Daniel, I. Chechuro, S. Verhees"
date: "20.02.2018 <br> Linguistic Convergence Lab Seminar"
output: word_document
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning= FALSE,
                      fig.width=10.4,
                      fig.height=7)
library(tidyverse)
library(lingtypology)
library(phangorn)
library(knitr)
```

# Similarity matrices

**Similarity matrix SV style**

Each meaning can correspond to one or more words in a given language. Each distinct lexeme gets its own row; languages are marked 1 if the lexeme/root is represented, 0 if this is not the case. Similarity is measured by the number of times languages share a 1 on the same row.


**Similarity matrix MD style**

Each meaning can correspond to one or more words in a given language. Each distinct lexeme is assigned a number; a distance matrix is created from the number of times languages share the same lexeme (i.e. share 1 or more numbers on a row) / number of meanings. 


# Visualizations

**Rutul area (SV style)**

1.
```{r}

Rutul_SV <- read.csv("Rutul_SV.csv", header = TRUE, sep = ";", na.strings = "NA")
Rutul_SV <- as.dist(Rutul_SV)
nnet <- neighborNet(Rutul_SV)
plot(nnet, "2D")

```


**Rutul area (MD style)**

2.
```{r}

Rutul_MD <- read.csv("Rutul_MD.csv", header = TRUE, sep = ";", na.strings = "NA")
Rutul_MD <- as.dist(Rutul_MD)
nnet <- neighborNet(Rutul_MD)
plot(nnet, "2D")

```

### Without "stop words"

If a meaning has a different lexeme in each language, remove it.
The idea is to measure distance based on loans only.


**Rutul area (SV style)**

3.
```{r}

Rutul_No_Stopwords_SV <- read.csv("Rutul_No_Stopwords_SV.csv", header = TRUE, sep = ";", na.strings = "NA")
Rutul_No_Stopwords_SV <- as.dist(Rutul_No_Stopwords_SV)
nnet <- neighborNet(Rutul_No_Stopwords_SV)
plot(nnet, "2D")

```


**Rutul area distance matrix (MD style)**

4.
```{r}

Rutul_No_Stopwords_MD <- read.csv("Rutul_No_Stopwords_MD.csv", header = TRUE, sep = ";", na.strings = "NA")
Rutul_No_Stopwords_MD <- as.dist(Rutul_No_Stopwords_MD)
nnet <- neighborNet(Rutul_No_Stopwords_MD)
plot(nnet, "2D")

```

### Distance between major languages
Based on a larger subset of the WOLD questionnaire.
Data gathered from dictionaries.
Visualizations based on the same variables as the Rutul data.


**Major languages (SV style)**

4.
```{r}

Big_LGS_SV <- read.csv("Big_LGS_SV.csv", header = TRUE, sep = ";", na.strings = "NA")
Big_LGS_SV <- as.dist(Big_LGS_SV)
nnet <- neighborNet(Big_LGS_SV)
plot(nnet, "2D")

```


** Major languages (MD style)**

5.
```{r}

Big_LGS_MD <- read.csv("Big_LGS_MD.csv", header = TRUE, sep = ";", na.strings = "NA")
Big_LGS_MD <- as.dist(Big_LGS_MD)
nnet <- neighborNet(Big_LGS_MD)
plot(nnet, "2D")

```

### Without "stop words"
If a meaning has a different lexeme in each language, remove it.
The idea is to measure distance based on loans only.


**Major languages (SV style)**

6.
```{r}

Big_LGS_No_Stopwords_SV <- read.csv("Big_LGS_No_Stopwords_SV.csv", header = TRUE, sep = ";", na.strings = "NA")
Big_LGS_SV <- as.dist(Big_LGS_No_Stopwords_SV)
nnet <- neighborNet(Big_LGS_No_Stopwords_SV)
plot(nnet, "2D")

```


**Major languages (MD style)**

7.
```{r}

Big_LGS_No_Stopwords_MD <- read.csv("Big_LGS_No_Stopwords_MD.csv", header = TRUE, sep = ";", na.strings = "NA")
Big_LGS_MD <- as.dist(Big_LGS_No_Stopwords_MD)
nnet <- neighborNet(Big_LGS_No_Stopwords_MD)
plot(nnet, "2D")

```

## Avar area (MD style)

8.
```{r}

avar <- read.csv("avar_dist.csv", header = TRUE, sep = ";", na.strings = "NA")
avar <- as.dist(avar)
nnet <- neighborNet(avar)
plot(nnet, "2D")

```

